---
title: "cds_biomarker"
output:
  github_document:
  html_notebook:
    theme: united
  html_document:
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = T)

library(tidyverse)
library(here)

source(here("R/load.R"))
source(here("R/linear_association.R"))
```

cds_biomarker is an R toolkit for biomarker analysis. It includes helpful functions and templates for standard analyses.

## Install

To install cds_biomarker clone this repository then run this command:

```{r, eval = FALSE}
install.packages("PATH_TO_LIBRARY/cds_biomarker", repos = NULL, type = "source")
```

## Templates

The [templates](https://github.com/broadinstitute/cds_biomarker/tree/master/templates) directory contains templates for standard analyses.

* [multi_profile_comparision](https://github.com/broadinstitute/cds_biomarker/tree/master/templates) compares biomarkers for multiple response profiles e.g drug and genetic or multiple drugs.

## Linear associations

`lin_associations()` can be used to calcualte the linear associations between a vector of responses and a matrix of biomarker features.

```{r results='hide'}
omics <- load_omics()
A <- omics[["RNA expression"]]
y <- load_achilles("PAX8")
```

```{r}
cls <- intersect(names(y),rownames(A))
lin_associations(A[cls,],y[cls]) %>% arrange(p.val) %>% head(5)
```
