---
title: "cdsr_biomarker"
output:
  github_document:
  html_notebook:
    theme: united
  html_document:
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = T)

library(tidyverse)
library(here)
library(useful)

source(here("R/biomarkers.R"))
```

cdsr_biomarker is an R toolkit for biomarker analysis. It includes helpful functions and standard reports.

## Install

To install cdsr_biomarker clone this repository then run this command:

```{r, eval = FALSE}
install.packages("PATH_TO_LIBRARY/cdsr_biomarker", repos = NULL, type = "source")
```

## Reports

The [reports](https://github.com/broadinstitute/cdsr_biomarker/tree/master/reports) directory contains standard biomarker reports.

* [multi_profile_biomarker_report](https://github.com/broadinstitute/cdsr_biomarker/tree/master/reports/multi_profile_biomarker_report.Rmd) compares biomarkers for multiple response profiles e.g drug and genetic or multiple drugs.

There are wrapper functions in cdsr_biomarker to automaticaly genenarate these reports. Here is an example:

1. Make a cell line by perturbation response matrix Y. Here I am using Achilles data for EGFR and PRISM data for a few EGFR inhibitors.

```{r include = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r eval = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r}
corner(Y)
```

2. Make a meta data table which will be displayed in the report

```{r}
meta_data <- list(perturbation = colnames(Y), type = c("CRISPR","Drug","Drug","Drug")) %>% as_tibble()
meta_data
```

3. Call the generate report function for the report you want. You will need to give it a file path to save the results to.

```{r eval = F}
generate_multi_profile_biomarker_report("~/Desktop/example/","example_title",Y,meta_data)
```

If you already have the biomarker results file and just want to generate the report you can do it like this:

```{r eval = F}
generate_multi_profile_biomarker_report("~/Desktop/example/","example_title")
```

