---
title: "cdsrbiomarker"
output:
  github_document:
  html_notebook:
    theme: united
  html_document:
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = T)

library(tidyverse)
library(useful)
library(taigr)
library(cdsrbiomarker)
```

cdsrbiomarker is an R toolkit for biomarker analysis. It includes helpful functions and standard reports.

## Install

```{r, eval = FALSE}
library(devtools)
devtools::install_github("broadinstitute/cdsr_biomarker")
```

The package can then be loaded by calling

```{r, eval=FALSE}
library(cdsrbiomarker)
```

## Taiga

Many of the function in the package use data which is stored on taiga. If you are at the Broad you can install taigr the taiga client for R by following the instructions [here](https://github.com/broadinstitute/taigr).

## Biomarker functions

These functions are used to analyze potential biomarkers based on a response vector and a feature matrix.

### discrete_test

Compares binary features, such as lineage and mutation, running a t-test on the difference in mean response between cell lines with the feature and without it. Run on response vector `y` and feature matrix `X`

```{r, eval=FALSE}
cdsrbiomarker::discrete_test(X, y)
```

### lin_associations

Compares continuous features, such as gene expression, calculating correlations between response and each feature. Run on feature matrix `A`, response vector `y`, and an optional matrix of confounders `W`. Other parameters can also be tuned and are explained in the function documentation.

```{r, eval=FALSE}
cdsrbiomarker::lin_associations(A, y, W=NULL)
```

### random_forest

Fits a random forest to a feature matrix `X` and a response vector `y` returning estimates of variable importance for each feature, as well as model level statistics such as R-squared. Other parameters can also be tuned and are explained in the function documentation.

```{r, eval=FALSE}
cdsrbiomarker::random_forest(X, y)
```

## Reports

The [reports](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports) directory contains standard biomarker reports.

* [single_profile_biomarker_report](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports/single_profile_biomarker_report.Rmd) generates a biomarker report for a single response profile.

* [multi_profile_biomarker_report](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports/multi_profile_biomarker_report.Rmd) compares biomarkers for multiple response profiles e.g drug and genetic or multiple drugs.

There are wrapper functions in cdsrbiomarker to automaticaly genenarate these reports. Here is an example using Achilles data for EGFR and PRISM data for a few EGFR inhibitors:

1. Make a cell line by perturbation response matrix Y.

```{r include = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r eval = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r}
corner(Y)
```

2. Make a meta data table which will be displayed in the report.

```{r}
meta_data <- list(perturbation = colnames(Y), type = c("CRISPR","Drug","Drug","Drug")) %>% as_tibble()
meta_data
```

3. Call the generate report function for the report you want. You will need to give it a file path to save the results to.

```{r eval = F}
cdsrbiomarker::generate_multi_profile_biomarker_report("~/Desktop/example/","example_title",Y,meta_data)
```

4. If you already have the biomarker results files and just want to generate the report you can do it like this.

```{r eval = F}
cdsrbiomarker::generate_multi_profile_biomarker_report("~/Desktop/example","example")
```

