---
title: "cdsrbiomarker"
output:
  github_document:
  html_notebook:
    theme: united
  html_document:
    theme: united
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = T)

library(tidyverse)
library(useful)
library(taigr)
library(cdsrbiomarker)
```

cdsrbiomarker is an R toolkit for biomarker analysis. It includes helpful functions and standard reports.

## Install

```{r, eval = FALSE}
library(devtools)
devtools::install_github("broadinstitute/cdsr_models")
devtools::install_github("broadinstitute/cdsr_biomarker")
```

The package can then be loaded by calling

```{r, eval=FALSE}
library(cdsrbiomarker)
```

## Taiga

Many of the functions in the package use data which is stored on taiga. If you are at the Broad you can install taigr the taiga client for R by following the instructions [here](https://github.com/broadinstitute/taigr).

## Functions

### get_biomarkers

`get_biomarkers` is a wrapper around the `discrete_test`, `lin_associations`, and `random_forest` functions in `cdsrmodels`. It takes a response vector `y` and then loads features matricies from Taiga. The appropriate modeling functions are applied to each feature matrix and the results are returned and saved in the `out_path` directory if it is specified.

```{r, eval=FALSE}
biomarker_results <- get_biomarkers(y, p_cutoff=0.1, out_path="~/Desktop/example")
```

## Reports

The [reports](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports) directory contains standard biomarker reports.

* [single_profile_biomarker_report](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports/single_profile_biomarker_report.Rmd) generates a biomarker report for a single response profile.

* [multi_profile_biomarker_report](https://github.com/broadinstitute/cdsr_biomarker/tree/master/inst/reports/multi_profile_biomarker_report.Rmd) compares biomarkers for multiple response profiles e.g drug and genetic or multiple drugs.

There are wrapper functions in cdsrbiomarker to automaticaly genenarate these reports. Here is an example using Achilles data for EGFR and PRISM data for a few EGFR inhibitors:

1. Make a cell line by perturbation response matrix Y.

```{r include = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r eval = F}
gene_effect <- load.from.taiga(data.name='depmap-a0ab',data.file='Achilles_gene_effect')[,"EGFR (1956)"] %>% 
  enframe(name = "arxspan_id",value = "xpr_egfr")
auc <- load.from.taiga(data.name='secondary-screen-15e6', data.file='secondary_merged_drc_parameters') %>% 
  filter(repurposing_name %in% c("erlotinib","gefitinib","lapatinib")) %>% 
  select(auc,repurposing_name,arxspan_id) %>% 
  spread(key = "repurposing_name",value = "auc")
Y <- full_join(gene_effect,auc, by = "arxspan_id") %>% column_to_rownames(var = "arxspan_id") %>% as.matrix()
```

```{r}
corner(Y)
```

2. Make a meta data table which will be displayed in the report.

```{r}
meta_data <- list(perturbation = colnames(Y), type = c("CRISPR","Drug","Drug","Drug")) %>% as_tibble()
meta_data
```

3. Call the generate report function for the report you want. You will need to give it a file path to save the results to.

```{r eval = F}
cdsrbiomarker::generate_multi_profile_biomarker_report("~/Desktop/example/","example_title",Y,meta_data)
```

4. If you already have the biomarker results files and just want to generate the report you can do it like this.

```{r eval = F}
cdsrbiomarker::generate_multi_profile_biomarker_report("~/Desktop/example","example")
```

