---
output:
  html_document:
  github_document: 
  html_notebook:
params:
  in_path: NULL
  title: "TITLE"
---

---
title: `r params$title`
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, echo=FALSE, include=FALSE}
library(taigr)
library(htmltools)
library(DT)
library(plotly)
library(crosstalk)
library(here)
library(cowplot)
library(tidyverse)
library(heatmaply)
```

```{r logo, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = F)
htmltools::img(src = knitr::image_uri(system.file("reports","CDS_logo.png",package = "cdsrbiomarker")), 
               alt = 'logo', 
               width="150px", height="170px",
               style = 'position:absolute; top:10px; left:75%')
```

```{r load data,include = F}
if (is.null(params$in_path)) {stop("You must specify the results directory")}
df <- read_csv(str_c(params$in_path,"/data.csv"))
y <- as.matrix(df[,-1])
rownames(y) <- df[[1]]
meta_data <- read_csv(str_c(params$in_path,"/meta_data.csv"))
disc_table <- read_csv(str_c(params$in_path,"/discrete_table.csv"))
lin_table <- read_csv(str_c(params$in_path,"/lin_association_table.csv"))
rf_table <- read_csv(str_c(params$in_path,"/rf_table.csv"))
```

This report presents biomarker analysis results for one or more response profiles. <br/> 
To navigate through this report click on the blue tabs under each heading.

# Response profiles {.tabset}

These tabs provide information about the response profiles included in this report.

## Info

```{r echo = FALSE}
datatable(meta_data, style="bootstrap", width="100%",options = list(lengthChange = FALSE))
```

## Correlation

```{r echo = FALSE,fig.width = 6, fig.height= 5,out.width = "60%"}
cor <- cor(y, use = "complete.obs")
diag(cor) <- NA
p <- plot_ly(
    x = colnames(cor), y = rownames(cor),
    z = cor, type = "heatmap"
)
p %>% colorbar(title = "Correlation") %>% layout(xaxis = list(tickangle = 90)) %>% config(displayModeBar = F)
```

# Discrete test {.tabset}

These tabs presents results for discrete features. The **effect size** is the difference between each feature and 
all others and the **q-value** is a t-test corrected for multiple hypotheses testing.  

```{r discrete test, results = 'asis', echo = FALSE}
for (cur_feature_type in disc_table$feature_type %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_type))
  for (cur_pert in disc_table$pert %>% unique()) {
    cat(sprintf('\n\n### %s \n\n', cur_pert))
    cat('\n\n You can scroll through the table to select features to 
      highlight in the volcano plot or select points in the volcano plot to view 
      them in the table.\n\n')
    
    df <- disc_table %>% 
      dplyr::filter(feature_type == cur_feature_type & pert == cur_pert & q.value < .1) %>%
      arrange(q.value) %>% head(2000) %>% 
      transmute(feature = feature,`effect size` = round(effect_size,4),
                `-log10(q-value)` = round(-log10(q.value),4),
                color = q.value < .01)
    
    colors <- df %>% arrange(color) %>% 
      mutate(color = ifelse(color,"darkorange","darkgrey")) %>% .[["color"]] %>% unique()

    sd <- highlight_key(select(df,-color),~feature)
    
    updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1,y = 1.15,
                             buttons = list(
      list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
          
    p <- plot_ly(type = "scatter", data = sd, x = ~`effect size`, y = ~`-log10(q-value)`,
               text = ~feature, hoverinfo = "text", color = df$color, 
               colors = colors,
               marker = list(line = list(width = 0))) %>% 
      layout(xaxis = list(title = "effect size"), yaxis = list(title = "-log10(q-value)"),
             showlegend = FALSE,updatemenus = updatemenus) %>% config(displayModeBar = F)
    
     p2 <-bscols(
      ggplotly(p) %>%
        highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                  off = "plotly_relayout",opacityDim = 1),
      datatable(sd, style="bootstrap", width="100%",
                options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
      widths = c(6,6))
     
     c('\n\n')
     cat(htmltools::knit_print.shiny.tag(p2))
     #cat(htmltools::renderTags(p2)$html)
     c('\n\n')
  }
  cat('\n\n### shared \n\n')
  cat('\n\n To generate this plot biomarkers for each response profile are ranked by q-value. The 50 biomarkers with the lowest average rank are
      across the response profiles are shown. \n\n')
  ranks <- disc_table %>% filter(feature_type == cur_feature_type) %>% dplyr::group_by(pert) %>% 
    dplyr::mutate(rank = dense_rank(q.value)) %>% select(feature,rank,pert) %>% 
    spread(key = "pert",value = "rank") %>% 
    tibble::column_to_rownames(var = "feature") %>% as.matrix()
  ranks[is.na(ranks)] <- max(ranks,na.rm = T)
  top_shared_biomarkers <- ranks %>% rowMeans() %>% sort() %>% head(50) %>% names()
  fig <- heatmaply(t(ranks[top_shared_biomarkers,]),dendrogram = "none",colors = viridis(n = 256, direction = -1),
                   column_text_angle = 90,na.rm = FALSE,key.title = "rank") %>% config(displayModeBar = F)
  c('\n\n')
  cat(htmltools::knit_print.shiny.tag(fig))
  c('\n\n')
}
```

# Linear association {.tabset}

These tabs presents results for continuous features. The **z-score** and **q-value** are calculated using linear modeling approach. 
Specifically, the **z-score** is the mean/standard deviation of the moderated effect size based on adaptive shrinkage.

```{r linear association, results = 'asis', echo = FALSE}
for (cur_feature_type in lin_table$feature_type %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_type))
  for (cur_pert in lin_table$pert %>% unique()) {
    cat(sprintf('\n\n### %s \n\n', cur_pert))
    cat('\n\n You can scroll through the table to select features to 
      highlight in the volcano plot or select points in the volcano plot to view 
      them in the table.\n\n')
    
    df <- lin_table %>% 
      dplyr::filter(feature_type == cur_feature_type & pert == cur_pert & qvalue < .1) %>%
      arrange(qvalue) %>% head(2000) %>% 
      transmute(feature = str_sub(feature,0,20),`z-score` = round(z.score,4),
                `-log10(q-value)` = round(-log10(qvalue),4),
                color = qvalue < .01)

    sd <- highlight_key(select(df,-color),~feature)
    
    updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1,y = 1.15,
                             buttons = list(
      list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
    
    colors <- df %>% arrange(color) %>% 
      mutate(color = ifelse(color,"darkorange","darkgrey")) %>% .[["color"]] %>% unique()
    
    p <- plot_ly(type = "scatter", data = sd, x = ~`z-score`, y = ~`-log10(q-value)`,
               text = ~feature, hoverinfo = "text", color = df$color, 
               colors = colors,
               marker = list(line = list(width = 0))) %>% 
      layout(xaxis = list(title = "z-score"), yaxis = list(title = "-log10(q-value)"),
             showlegend = FALSE,updatemenus = updatemenus) %>% config(displayModeBar = F)
    
     p2 <-bscols(
      ggplotly(p) %>%
        highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                  off = "plotly_relayout",opacityDim = 1),
      datatable(sd, style="bootstrap", width="100%",
                options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
      widths = c(6,6))
     
     c('\n\n')
     cat(htmltools::knit_print.shiny.tag(p2))
     c('\n\n')
  }
  cat('\n\n### shared \n\n')
  cat('\n\n To generate this plot biomarkers for each response profile are ranked by q-value. The 50 biomarkers with the lowest average rank are
      across the response profiles are shown. \n\n')
  ranks <- lin_table %>% filter(feature_type == cur_feature_type) %>% dplyr::group_by(pert) %>% 
    dplyr::mutate(rank = rank(qvalue)) %>% select(feature,rank,pert) %>% 
    spread(key = "pert",value = "rank") %>% 
    tibble::column_to_rownames(var = "feature") %>% as.matrix()
  ranks[is.na(ranks)] <- max(ranks,na.rm = T)
  ranks[ranks > 500] <- 500
  top_shared_biomarkers <- ranks %>% rowMeans() %>% sort() %>% head(50) %>% names()
  fig <- heatmaply(t(ranks[top_shared_biomarkers,]),dendrogram = "none",colors = viridis(n = 256, direction = -1),
                   column_text_angle = 90,na.rm = FALSE,key.title = "rank") %>% config(displayModeBar = F)
  c('\n\n')
  cat(htmltools::knit_print.shiny.tag(fig))
  c('\n\n')
}
```

# Random forest {.tabset}

These tabs presents random forest results for different feature sets. **Feature importance** is the importance of the feature in the random forstest and **stability** is the stability of that feature are folds in 10-fold cross validation.

```{r random forest, results = 'asis', echo = FALSE,fig.width = 7, fig.height= 5,out.width = "70%"}
cat('\n\n## model accuracy\n\n')
cat('\n\n This plot shows the pearson correlation between the random forest predictions 
    and the measured values.\n\n')

rf_table %>% group_by(pert,feature_set) %>% 
  dplyr::summarize(pearson = first(PearsonScore)) %>% 
  ggplot(aes(pert,pearson,fill = feature_set)) + geom_col(position = "dodge") + 
  theme_cowplot(10) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "perturbation", fill = "feature set") 

for (cur_feature_set in rf_table$feature_set %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_set))
  for (cur_pert in rf_table$pert %>% unique()) {
    cat(sprintf('\n\n### %s \n\n', cur_pert))
    cat('\n\n You can scroll through the table to select features to 
      highlight in the plot or select points in the plot to view 
      them in the table.\n\n')
    
    df <- rf_table %>% 
      dplyr::filter(feature_set == cur_feature_set & pert == cur_pert) %>% 
      transmute(feature = feature,`feature importance` = round(RF.imp.mean,4),
                stability = round(RF.imp.stability,4),
                color = str_split_fixed(feature,pattern = "_",n = 2)[,1])

    sd <- highlight_key(select(df,-color),~feature)
    
    updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1.2,y = 1.15,
                             buttons = list(
      list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
          
    p <- plot_ly(type = "scatter", data = sd, x = ~`feature importance`, y = ~stability,
               text = ~feature, hoverinfo = "text", color = df$color, colors = "Set1",
               marker = list(line = list(width = 0))) %>% 
      layout(xaxis = list(title = "feature importance"), yaxis = list(title = "stability"),
             showlegend = TRUE,updatemenus = updatemenus) %>% config(displayModeBar = F)
    
     p2 <-bscols(
      ggplotly(p) %>%
        highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                  off = "plotly_relayout",opacityDim = 1),
      datatable(sd, style="bootstrap", width="100%",
                options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
      widths = c(6,6))
     
     c('\n\n')
     cat(htmltools::knit_print.shiny.tag(p2))
     c('\n\n')
  }
  cat('\n\n### shared \n\n')
  cat('\n\n To generate this plot biomarkers for each response profile are ranked by q-value. The 50 biomarkers with the lowest average rank are
      across the response profiles are shown. \n\n')
  ranks <- rf_table %>% filter(feature_set == cur_feature_set) %>% dplyr::group_by(pert) %>% 
    dplyr::mutate(rank = rank(-RF.imp.mean)) %>% select(feature,rank,pert) %>% 
    spread(key = "pert",value = "rank") %>% 
    tibble::column_to_rownames(var = "feature") %>% as.matrix()
  ranks[is.na(ranks)] <- max(ranks,na.rm = T)
  top_shared_biomarkers <- ranks %>% rowMeans() %>% sort() %>% head(50) %>% names()
  fig <- heatmaply(t(ranks[top_shared_biomarkers,]),dendrogram = "none",colors = viridis(n = 256, direction = -1),
                   column_text_angle = 90,na.rm = FALSE,key.title = "rank") %>% config(displayModeBar = F)
  c('\n\n')
  cat(htmltools::knit_print.shiny.tag(fig))
  c('\n\n')
}
```


