---
output:
  html_document:
  github_document: 
  html_notebook:
params:
  in_path: NULL
  title: "TITLE"
---

---
title: `r params$title`
date: "`r format(Sys.time(), '%d %B %Y')`"
---

```{r setup, echo=FALSE, include=FALSE}
library(taigr)
library(htmltools)
library(DT)
library(plotly)
library(crosstalk)
library(here)
library(cowplot)
library(tidyverse)
library(heatmaply)
```

```{r logo, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 7, cache = F)
htmltools::img(src = knitr::image_uri(system.file("reports","CDS_logo.png",package = "cdsrbiomarker")), 
               alt = 'logo', 
               width="150px", height="170px",
               style = 'position:absolute; top:10px; left:75%')
```

```{r, include = F}
bscols(plot_ly(),datatable(tibble(),style="bootstrap"))
```

```{r load data,include = F}
if (is.null(params$in_path)) {stop("You must specify the results directory")}
df <- read_csv(str_c(params$in_path,"/data.csv"))
y <- as.matrix(df[,-1])
rownames(y) <- df[[1]]
disc_table <- read_csv(str_c(params$in_path,"/discrete_table.csv"))
lin_table <- read_csv(str_c(params$in_path,"/lin_association_table.csv"))
rf_table <- read_csv(str_c(params$in_path,"/rf_table.csv"))
```

This report presents biomarker analysis results for a drug or genetic response profile. <br/> 
To navigate through this report click on the blue tabs under each heading.

# Info

```{r echo = FALSE}
if(file.exists(str_c(params$in_path,"/meta_data.csv"))){
  meta_data <- read_csv(str_c(params$in_path,"/meta_data.csv"))
  datatable(meta_data, style="bootstrap", width="100%",options = list(lengthChange = FALSE))
}
```

# Discrete test {.tabset .tabset-pills}

These tabs presents results for discrete features. The **effect size** is the difference between each feature and 
all others and the **q-value** is a t-test corrected for multiple hypotheses testing.  

```{r discrete test, results = 'asis', echo = FALSE}
for (cur_feature_type in disc_table$feature_type %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_type))
  cat('\n\n You can scroll through the table to select features to 
    highlight in the volcano plot or select points in the volcano plot to view 
    them in the table.\n\n')
  
  df <- disc_table %>% 
    dplyr::filter(feature_type == cur_feature_type & q.value < .1) %>%
    arrange(q.value) %>% head(2000) %>% 
    transmute(feature = feature,`effect size` = round(effect_size,4),
              `-log10(q-value)` = round(-log10(q.value),4),
              color = q.value < .01)
  
  colors <- df %>% arrange(color) %>% 
    mutate(color = ifelse(color,"darkorange","darkgrey")) %>% .[["color"]] %>% unique()

  sd <- highlight_key(select(df,-color),~feature)
  
  updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1,y = 1.15,
                           buttons = list(
    list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
        
  p <- plot_ly(type = "scatter", data = sd, x = ~`effect size`, y = ~`-log10(q-value)`,
             text = ~feature, hoverinfo = "text", color = df$color, 
             colors = colors,
             marker = list(line = list(width = 0))) %>% 
    layout(xaxis = list(title = "effect size"), yaxis = list(title = "-log10(q-value)"),
           showlegend = FALSE,updatemenus = updatemenus) %>% config(displayModeBar = F)
  
   p2 <-bscols(
    ggplotly(p) %>%
      highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                off = "plotly_relayout",opacityDim = 1),
    datatable(sd, style="bootstrap", width="100%",
              options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
    widths = c(6,6))
   
   c('\n\n')
   cat(htmltools::knit_print.shiny.tag(p2))
   c('\n\n')
}
```

# Linear association {.tabset .tabset-pills}

These tabs presents results for continuous features. The **effect size** and **q-value** are calculated using linear modeling approach. 
Specifically, the **effect size** is the mean moderated effect size based on adaptive shrinkage.

```{r linear association, results = 'asis', echo = FALSE}
for (cur_feature_type in lin_table$feature_type %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_type))
  cat('\n\n You can scroll through the table to select features to 
    highlight in the volcano plot or select points in the volcano plot to view 
    them in the table.\n\n')
  
  df <- lin_table %>% 
    dplyr::filter(feature_type == cur_feature_type & qvalue < .1) %>%
    arrange(qvalue) %>% head(2000) %>% 
    transmute(feature = str_sub(ind.var,0,20),`effect size` = round(PosteriorMean,4),
              `-log10(q-value)` = round(-log10(qvalue),4),
              color = qvalue < .01)

  sd <- highlight_key(select(df,-color),~feature)
  
  updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1,y = 1.15,
                           buttons = list(
    list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
  
  colors <- df %>% arrange(color) %>% 
    mutate(color = ifelse(color,"darkorange","darkgrey")) %>% .[["color"]] %>% unique()
  
  p <- plot_ly(type = "scatter", data = sd, x = ~`effect size`, y = ~`-log10(q-value)`,
             text = ~feature, hoverinfo = "text", color = df$color, 
             colors = colors,
             marker = list(line = list(width = 0))) %>% 
    layout(xaxis = list(title = "effect size"), yaxis = list(title = "-log10(q-value)"),
           showlegend = FALSE,updatemenus = updatemenus) %>% config(displayModeBar = F)
  
   p2 <-bscols(
    ggplotly(p) %>%
      highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                off = "plotly_relayout",opacityDim = 1),
    datatable(sd, style="bootstrap", width="100%",
              options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
    widths = c(6,6))
   
   c('\n\n')
   cat(htmltools::knit_print.shiny.tag(p2))
   c('\n\n')
}
```

# Random forest {.tabset .tabset-pills}

These tabs presents random forest results for different feature sets. **Feature importance** is the importance of the feature in the random forstest and **stability** is the stability of that feature are folds in 10-fold cross validation.

```{r random forest, results = 'asis', echo = FALSE,fig.width = 7, fig.height= 5,out.width = "70%"}
cat('\n\n## model accuracy\n\n')
cat('\n\n This plot shows the pearson correlation between the random forest predictions 
    and the measured values.\n\n')

rf_table %>% group_by(feature_set) %>% 
  dplyr::summarize(pearson = first(PearsonScore)) %>% 
  ggplot(aes(feature_set,pearson,fill = feature_set)) + geom_col(position = "dodge") + 
  theme_cowplot(10) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "feature set") + guides(fill = F)

for (cur_feature_set in rf_table$feature_set %>% unique()) {
  cat(sprintf('\n\n## %s {.tabset .tabset-pills}\n\n', cur_feature_set))
  cat('\n\n You can scroll through the table to select features to 
    highlight in the plot or select points in the plot to view 
    them in the table.\n\n')
  
  df <- rf_table %>% 
    dplyr::filter(feature_set == cur_feature_set) %>% 
    transmute(feature = feature,`feature importance` = round(RF.imp.mean,4),
              stability = round(RF.imp.stability,4),
              color = str_split_fixed(feature,pattern = "_",n = 2)[,1])

  sd <- highlight_key(select(df,-color),~feature)
  
  updatemenus <- list(list(showactive = FALSE,type = 'buttons',x = 1.2,y = 1.15,
                           buttons = list(
    list(label = "Reset",method = "relayout",args = list(list(shapes = c()))))))
        
  p <- plot_ly(type = "scatter", data = sd, x = ~`feature importance`, y = ~stability,
             text = ~feature, hoverinfo = "text", color = df$color, colors = "Set1",
             marker = list(line = list(width = 0))) %>% 
    layout(xaxis = list(title = "feature importance"), yaxis = list(title = "stability"),
           showlegend = TRUE,updatemenus = updatemenus) %>% config(displayModeBar = F)
  
   p2 <-bscols(
    ggplotly(p) %>%
      highlight(color = rgb(0,0.4588,0.6901),on = "plotly_click",
                off = "plotly_relayout",opacityDim = 1),
    datatable(sd, style="bootstrap", width="100%",
              options = list(lengthChange = FALSE,scrollY = "300px",paging = FALSE)),
    widths = c(6,6))
   
   c('\n\n')
   cat(htmltools::knit_print.shiny.tag(p2))
   c('\n\n')
}
```


